A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-11-13 docs/results_JAMwithNFW_v3.01.tex
Included corrections in abstract and introduction.

\paragraph{Modelling and priors.} The full set of fit parameters is $(\Upsilon_\text{I,*},r_s,v_{200},\beta_z)$, where $\beta_z$ is the constant velocity anisotropy parameter (see Section \ref{sec:model_JAM}). We will investigate this parameter space with a MCMC\footnote{The Python code package for \emph{emcee}, a Monte Carlo Markov Chain implementation by \citet{emcee} is available online at \url{http://dan.iel.fm/emcee/current/}. The version from October 2013 was used in this work.} \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For a stellar population with a Chabrier IMF they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate, we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Equation \ref{eq:Maccio08} as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{eqnarray*}
&&p(\Upsilon_\text{I,*},r_s,v_{200},\beta) \\
&&= \frac{1}{\ln\left( 10 r_s\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{eqnarray*}
where the factor $1/\ln\left( 10 r_s\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ to $(v_{200},r_s)$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5$ arcsec, approximately within the effective half-light radius $R_\text{eff} = 2.6$ arcsec. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit. [TO DO: Didn't include the results from the Moster 2010 relation, because they where quite close to the Dutton 2010 relation anyway. Or should I?]

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $r_s$) than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies: The sample with the highest probability is pegged at the lower limit of the unifrom prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. There are slight covariances between $\Upsilon_\text{I,*}$,  $r_s$ and $\beta_z$: The smaller the dark matter contribution in the center (i.e. the larger $\Upsilon_\text{I,*}$), the less concentrated is the halo (i.e. the larger $r_s$), the more velocity anisotropy is needed to reproduce the central dip. As the effect of $v_{200}$ is mostly at larger radii, this parameter doesn't show any covariances - but is also mostly constrained by the prior. The $v_\text{rms}$ curve corresponding to the mean best fit values from Table \ref{tab:modelB4_bestfit} is shown in the first panel of Figure \ref{fig:modelB4_vrot}. Figure \ref{fig:modelB4_models} illustrates the range of parameters fitting the data according to the full posterior probability function. The model fits the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and is also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6$ arcsec at all and seems to be therefore overall too massive.\\

@@

\begin{figure*}
\centering
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{fig/B4_rms_error_curves.ps}
  \caption{Comparison of the $v_\text{rms}$ data with the best fit JAM models.}
  \label{fig:modelB4_vrms}
\end{subfigure}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Circular velocity curve.}
  \label{fig:modelB4_vcirc}
\end{subfigure}%
\hspace{.02\textwidth}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_projmass.ps}
  \caption{Enclosed mass profile.}
  \label{fig:modelB4_enclMass}
\end{subfigure}
\caption{Best fit JAM model including a NFW halo and velocity anisotropy with parameters given in Table \ref{tab:modelB4_bestfit}. The 12 lines shown correspond to the 12 models randomly drawn from the posterior probability distribution and marked as grey diamonds in Figure \ref{fig:modelB4_triangle}. Overplotted are the Einstein radius (red dotted line) and the effective half-light radius (black dotted line). The blue dashed line marks the radius within which the data and model were fitted. \emph{Panel a)} shows the comparison of the symmetrized $v_\text{rms}$ data (solid blue points) with the best fit JAM models including a NFW halo (red solid lines). Also shown is the non-symmetrized data at larger radii (open blue dots) and an extrapolation of the best fit models, using the same model parameters but the I-band surface brightness MGE for the outer regions of J1331 derived from the Ellipse model (red dashed lines). \emph{Panel b)} shows the circular velocity curve of the total mass (green), and separately the contribution of the stellar mass (yellow, again generated from the MGE in Table \ref{tab:MGEF814W}) and dark matter (grey). [TO DO: maybe make one separate circular velocity plot, that also contains the mass-follows-light with Einstein M/L vcirc and compares it with Aaron's vcirc curve.] \emph{Panel c)} shows the corresponding enclosed mass profile. Overplotted is also the Einstein mass at the Einstein radius with a 10\% error, which was used in the fit as an additional constraint.}
\label{fig:modelB4_models}
\end{figure*}

\paragraph{Rotation curve.} Following the procedure in Section \ref{sec:model_JAM} we find the rotation curve from the best fit mean model in Table \ref{tab:modelB4_bestfit} by fitting the rotation parameter $\kappa'$ to the symmetrized $v_\text{rot}$ data within $R' = 3.5$ arcsec. The best fit with $\kappa' = 0.76$ is shown in the second panel of Figure \ref{fig:modelB4_vrot}. The third panel gives the dispersion that was calculated simply by $\sigma = \sqrt{v_\text{rms}^2 - v_\text{rot}^2}$. Our assumptions for $\kappa(R)$ nicely reproduce a $v_\text{rot}$ model with counter-rotating core. Although we fitted $v_\text{rot}$ only to the inner regions, the extrapolation to large radii fits the data also very well. While the dispersion $\sigma$ in the center fits by construction quite good, there is a huge discrepancy at large radii. The predicted dispersion is much larger than the data; we would expect the disk rotationally supported and therefore have a low velocity dispersion; especially dispersions as high as $\sim 200 km/s$ are more likely to be observed in the pressure supported bulges of galaxies. There \emph{might} be something unexpeted with the $\sigma$ measurements around $\sim 5$ arcsec, but at large radii the the best fit model NFW halo is simply too massive.

\begin{figure*}
\centering
\includegraphics[width=0.7\linewidth]{fig/B4_rms_rot_curves_best_model.ps}
\caption{Generating the rotation curve from the JAM model with NFW halo and constant velocity anisotropy with the mean parameters in Table \ref{tab:modelB4_bestfit} and with the best fit rotation parameter $\kappa' = 0.76$ in Equation \ref{eq:kappa_profile} (red solid lines). The second velocity moment in the first panel and the first velocity moment in the second panel with the additional fit parameter $\kappa'$ were fitted to the symmetrized data (solid blue points). The velocity dispersion is simply $\sigma = \sqrt{v_\text{rms}^2 - v_\text{rot}^2}$. At larger radii we compare the unsymmetrized data (open blue dots), the gas kinematics from \citet{SWELLSV} (green diamonds) and a JAM model using the same model parameters but the light distribution MGE generated from the Ellipse model in Figure \ref{fig:MGEouterRegions} as a prediction for the outer regions of J1331 (red dashed lines). The central regions are very well reproduced and we can also greatly predict the rotation curve at larger radii. Only at larger radii the $v_\text{rms}$ and $v_\text{rot}$ overestimate the measurements, probably due to a too massive NFW halo.}
\label{fig:modelB4_vrot}
\end{figure*}

>>>>

\paragraph{Modelling and priors.} The full set of fit parameters is $(\Upsilon_\text{I,*},r_s,v_{200},\beta_z)$, where $\beta_z$ is the constant velocity anisotropy parameter (see Section \ref{sec:model_JAM}). We will investigate this parameter space with a MCMC\footnote{The Python code package for \emph{emcee}, a Monte Carlo Markov Chain implementation by \citet{emcee} is available online at \url{http://dan.iel.fm/emcee/current/}. The version from October 2013 was used in this work.} \citep{emcee} and use priors for the halo parameters to guide the fit to a realistic NFW halo shape.
\\\citet{Dutton10} give a relation for halo vs. stellar mass for late-type galaxies. For a stellar population with a Chabrier IMF they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ \citet{Dutton10}. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate, we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by \citet{Dutton10}. The second error is the propagated error due to the uncertainty in the stellar mass. We use this as a rough estimate for the halo of J1331 and as Gaussian prior on $v_{200}$, 
\begin{equation*}
p(v_{200}) = \mathscr{N}(200 \text{km/s} ,40 \text{km/s}).
\end{equation*}
\\We also use the concentration vs. halo mass relation by \citet{Maccio08} in Equation \ref{eq:Maccio08} as a prior on the concentration, i.e.
\begin{equation*}
p(\log c_{200} \mid v_{200}) = \mathscr{N}\left(\langle \log c_{200} \rangle (M_{200}) \mid 0.105 \right).
\end{equation*}
\\For the velocity anisotropy parameter $\beta$ we will again employ a uniform prior 
\begin{equation*}
p(\beta) = \mathscr{U}(-0.5,+0.5)
\end{equation*}
to exclude very unrealistic anisotropies. The full prior used is then
\begin{eqnarray*}
&&p(\Upsilon_\text{I,*},r_s,v_{200},\beta) \\
&&= \frac{1}{\ln\left( 10 r_s\right)} p(\log c_{200} \mid v_{200}) \cdot p(v_{200}) \cdot p(\beta), 
\end{eqnarray*}
where the factor $1/\ln\left( 10 r_s\right)$ is the Jacobian of the transformation from the halo parameters $(v_{200},\log c_{200})$ to $(v_{200},r_s)$.
\\We restrict the $v_\text{rms}$ fit to a region $R' \lesssim 3.5''$, approximately within the effective half-light radius $R_\text{eff} = 2.6''$. We also include the Einstein mass $M(<R_\text{ein}$ with a 10\% error as an additional constraint in the fit. \Wilma{[TO DO: Didn't include the results from the Moster 2010 relation, because they where quite close to the Dutton 2010 relation anyway. Or should I?]}

\paragraph{Result.} Figure \ref{fig:modelB4_triangle} shows the posterior probability distribution of the fit sampled with an MCMC. Overplotted are also the priors used to constrain the NFW halo. The mean parameters are summarized in Table \ref{tab:modelB4_bestfit}. We find that the best fit NFW halo strives to be more massive and with a higher concentration (due to a smaller $r_s$) than  proposed by the prior. At the same time the model prefers again very negative velocity anisotropies: The sample with the highest probability is pegged at the lower limit of the unifrom prior. Both, the concentrated halo and the low anisotropy are needed to reproduce the central dip of the $v_\text{rms}$ curve. There are slight covariances between $\Upsilon_\text{I,*}$,  $r_s$ and $\beta_z$: The smaller the dark matter contribution in the center (i.e. the larger $\Upsilon_\text{I,*}$), the less concentrated is the halo (i.e. the larger $r_s$), the more velocity anisotropy is needed to reproduce the central dip. As the effect of $v_{200}$ is mostly at larger radii, this parameter doesn't show any covariances - but is also mostly constrained by the prior. The $v_\text{rms}$ curve corresponding to the mean best fit values from Table \ref{tab:modelB4_bestfit} is shown in the first panel of Figure \ref{fig:modelB4_vrot}. Figure \ref{fig:modelB4_models} illustrates the range of parameters fitting the data according to the full posterior probability function. The model fits the $v_\text{rms}$ data in the inner regions quite well (Figure \ref{fig:modelB4_vrms}) and is also consistent with the Einstein mass (see Figure \ref{fig:modelB4_enclMass}). The extrapolation of the model to larger radii however overestimates the data, does not exhibit a drop around $\sim 6''$ at all and seems to be therefore overall too massive.\\

@@

\begin{figure*}
\centering
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=0.8\linewidth]{fig/B4_rms_error_curves.ps}
  \caption{Comparison of the $v_\text{rms}$ data with the best fit JAM models.}
  \label{fig:modelB4_vrms}
\end{subfigure}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_vcirc.ps}
  \caption{Circular velocity curve.}
  \label{fig:modelB4_vcirc}
\end{subfigure}%
\hspace{.02\textwidth}
\begin{subfigure}{.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{fig/B4_jam_profiles_errors_short_projmass.ps}
  \caption{Enclosed mass profile.}
  \label{fig:modelB4_enclMass}
\end{subfigure}
\caption{Best fit JAM model including a NFW halo and velocity anisotropy with parameters given in Table \ref{tab:modelB4_bestfit}. The 12 lines shown correspond to the 12 models randomly drawn from the posterior probability distribution and marked as grey diamonds in Figure \ref{fig:modelB4_triangle}. Overplotted are the Einstein radius (red dotted line) and the effective half-light radius (black dotted line). The blue dashed line marks the radius within which the data and model were fitted. \emph{Panel a)} shows the comparison of the symmetrized $v_\text{rms}$ data (solid blue points) with the best fit JAM models including a NFW halo (red solid lines). Also shown is the non-symmetrized data at larger radii (open blue dots) and an extrapolation of the best fit models, using the same model parameters but the I-band surface brightness MGE for the outer regions of J1331 derived from the Ellipse model (red dashed lines). \emph{Panel b)} shows the circular velocity curve of the total mass (green), and separately the contribution of the stellar mass (yellow, again generated from the MGE in Table \ref{tab:MGEF814W}) and dark matter (grey). \Wilma{[TO DO: maybe make one separate circular velocity plot, that also contains the mass-follows-light with Einstein M/L vcirc and compares it with Aaron's vcirc curve.]} \emph{Panel c)} shows the corresponding enclosed mass profile. Overplotted is also the Einstein mass at the Einstein radius with a 10\% error, which was used in the fit as an additional constraint. \Wilma{[TO DO: Replace arcsec with '']}}
\label{fig:modelB4_models}
\end{figure*}

\paragraph{Rotation curve.} Following the procedure in Section \ref{sec:model_JAM} we find the rotation curve from the best fit mean model in Table \ref{tab:modelB4_bestfit} by fitting the rotation parameter $\kappa'$ to the symmetrized $v_\text{rot}$ data within $R' = 3.5''$. The best fit with $\kappa' = 0.76$ is shown in the second panel of Figure \ref{fig:modelB4_vrot}. The third panel gives the dispersion that was calculated simply by $\sigma = \sqrt{v_\text{rms}^2 - v_\text{rot}^2}$. Our assumptions for $\kappa(R)$ nicely reproduce a $v_\text{rot}$ model with counter-rotating core. Although we fitted $v_\text{rot}$ only to the inner regions, the extrapolation to large radii fits the data also very well. While the dispersion $\sigma$ in the center fits by construction quite good, there is a huge discrepancy at large radii. The predicted dispersion is much larger than the data; we would expect the disk rotationally supported and therefore have a low velocity dispersion; especially dispersions as high as $\sim 200 km/s$ are more likely to be observed in the pressure supported bulges of galaxies. There \emph{might} be something unexpeted with the $\sigma$ measurements around $\sim 5''$, but at large radii the the best fit model NFW halo is simply too massive.

\begin{figure*}
\centering
\includegraphics[width=0.7\linewidth]{fig/B4_rms_rot_curves_best_model.ps}
\caption{Generating the rotation curve from the JAM model with NFW halo and constant velocity anisotropy with the mean parameters in Table \ref{tab:modelB4_bestfit} and with the best fit rotation parameter $\kappa' = 0.76$ in Equation \ref{eq:kappa_profile} (red solid lines). The second velocity moment in the first panel and the first velocity moment in the second panel with the additional fit parameter $\kappa'$ were fitted to the symmetrized data (solid blue points). The velocity dispersion is simply $\sigma = \sqrt{v_\text{rms}^2 - v_\text{rot}^2}$. At larger radii we compare the unsymmetrized data (open blue dots), the gas kinematics from \citet{SWELLSV} (green diamonds) and a JAM model using the same model parameters but the light distribution MGE generated from the Ellipse model in Figure \ref{fig:MGEouterRegions} as a prediction for the outer regions of J1331 (red dashed lines). The central regions are very well reproduced and we can also greatly predict the rotation curve at larger radii. Only at larger radii the $v_\text{rms}$ and $v_\text{rot}$ overestimate the measurements, probably due to a too massive NFW halo. \Wilma{[TO DO: Replace arcsec with '']}}
\label{fig:modelB4_vrot}
\end{figure*}
